The ChEMBL database as Linked Open Data

### 2012-04-02 article.tex
Fixed column overflows

The full set of triples are available from http://semantics.bigcat.unimaas.nl/chembl12/.

@@

\begin{figure}[t]
\includegraphics[width=0.5\textwidth]{figs/relations}
\caption{The various resource types found in the ChEMBL triples. Some entities are subclasses
of common classes, while others are instances.}\label{f1}
\end{figure}

@@

\begin{small}
\begin{verbatim}
assay:a17
 a chembl:Assay ;
 chembl:extractedFrom
  res:r11347 ;
 chembl:hasAssayType chembl:ADMET ;
 chembl:hasConfScore "7"^^xsd:int ;
 chembl:hasDescription
  "Inhibition of cytochrome P450 progesterone 15-alpha hydroxylase" ;
 chembl:hasTarget target:t100122 .
\end{verbatim}
\end{small}

@@

\begin{figure}[t]
\includegraphics[width=0.5\textwidth]{figs/lodgraph}
\caption{The links out of the ChEMBL-RDF data into the Linked Open Data cloud.
Edges are labeled by the predicates making the links.}\label{2}
\end{figure}

@@

\begin{figure}[bt]
\includegraphics[width=0.5\textwidth]{snorql}
\caption{The SNORQL interface for browsing the SPARQL end point.}\label{f2}
\end{figure}

@@

The Linked Data server used by Bio2RDF has been reconfigured for ChEMBL to provide URL
based services for standard URI resolution, along with text and link searches
\footnote{\url{https://github.com/ansell/chembl-webapp}}. It proxies the standard item
URIs by translating URLs between those requested by users and the URIs that are available
in the databases. For example, if the ChEMBL web application is running on the users
local machine \footnote{\url{http://localhost:8080/chembl/}}, then a request for the
article with identifier ``a31863'' \footnote{\url{http://localhost:8080/chembl/article/a31863}}
will be resolved from the database using the full original URI
\footnote{\url{http://data.kasabi.com/dataset/chembl-rdf/13/activity/a31863}}. If the
user requested an RDF document, using content negotiation, the original URIs will be unchanged,
however, if the user requested an HTML document, the results will contain both the
original RDF triples, represented using RDFa, with links that resolve using the users local machine.

The links services enables the Chembl application to derive both forward links, originating in
ChEMBL \footnote{\url{http://localhost:8080/chembl/linkstonamespace/targetns/originalns:identifier}},
and backward links, originating in other databases, such as LODD, Bio2RDF
\footnote{\url{http://bio2rdf.org/linksns/targetns/originalns:identifier}}, or Chem2Bio2RDF.
These services are vital to efficiently navigate the Linked Data web, as it is both
impractical and inefficient to require users to crawl the web before they can discover all of the
relevant resources. These services are currently only supplied as web services from ChEMBL and
Bio2RDF, but it is hoped that similar services will be provided by other scientific Linked Data
providers in the future. Datasets that are available in SPARQL endpoints can be queried for
links efficiently using templated queries in the ChEMBL web application.

>>>>

The full set of triples are available from the website http://semantics.bigcat.unimaas.nl/chembl12/.

@@

\begin{figure}[t]
\includegraphics[width=0.45\textwidth]{figs/relations}
\caption{The various resource types found in the ChEMBL triples. Some entities are subclasses
of common classes, while others are instances.}\label{f1}
\end{figure}

@@

\begin{small}
\begin{verbatim}
assay:a17
 a chembl:Assay ;
 chembl:extractedFrom
  res:r11347 ;
 chembl:hasAssayType chembl:ADMET ;
 chembl:hasConfScore "7"^^xsd:int ;
 chembl:hasDescription
  "Inhibition of ... hydroxylase" ;
 chembl:hasTarget target:t100122 .
\end{verbatim}
\end{small}

@@

\begin{figure}[t]
\includegraphics[width=0.45\textwidth]{figs/lodgraph}
\caption{The links out of the ChEMBL-RDF data into the Linked Open Data cloud.
Edges are labeled by the predicates making the links.}\label{2}
\end{figure}

@@

\begin{figure}[bt]
\includegraphics[width=0.45\textwidth]{snorql}
\caption{The SNORQL interface for browsing the SPARQL end point.}\label{f2}
\end{figure}

@@

The Linked Data server used by Bio2RDF has been reconfigured for ChEMBL to provide URL
based services for standard URI resolution, along with text and link searches
\cite{WebAppGitHub}. It proxies the standard item
URIs by translating URLs between those requested by users and the URIs that are available
in the databases. For example, if the ChEMBL web application is running on the users
local machine, e.g. \url{http://localhost:8080/chembl/}, then a request for the
article with identifier ``a31863'', \url{http://localhost:8080/-chembl/article/a31863},
will be resolved from the database using the full original URI,
\url{http://data.kasabi.com/-dataset/chembl-rdf/13/activity/a31863}. If the
user requested an RDF document, using content negotiation, the original URIs will be unchanged,
however, if the user requested an HTML document, the results will contain both the
original RDF triples, represented using RDFa, with links that resolve using the users local machine.

The links services enables the ChEMBL application to derive both forward links, originating in
ChEMBL, e.g. \url{http://localhost:8080/chembl/linkstonamespace/-targetns/originalns:identifier},
and backward links, originating in other databases, such as LODD, Bio2RDF, e.g.
\url{http://bio2rdf.org/linksns/targetns/originalns:ident-ifier}, or Chem2Bio2RDF.
These services are vital to efficiently navigate the Linked Data web, as it is both
impractical and inefficient to require users to crawl the web before they can discover all of the
relevant resources. These services are currently only supplied as web services from ChEMBL and
Bio2RDF, but it is hoped that similar services will be provided by other scientific Linked Data
providers in the future. Datasets that are available in SPARQL endpoints can be queried for
links efficiently using templated queries in the ChEMBL web application.
